---
title: "Project 1 Interim"
author: "Kyle Kumbier"
date: "Jan 22, 2019"
output:
  beamer_presentation:
    colortheme: wolverine
    fonttheme: structurebold
    theme: AnnArbor
  ioslides_presentation: default
  slidy_presentation: default
fontsize: 20pt
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

setwd("~/WORKING_DIRECTORIES/biostat.699/project01")

library(readr)
library(dplyr)
dataset01 <- read.csv("dataset01.csv") # load cleaned version of thedata.csv
dataset02 <- read.csv("dataset02.csv") # load cleaned version of Copy+of+Elderly+Driver+Data.xlsx
dataset02 <- dataset02[,1:38] # has three unnecessary columns when loaded for some reason
colnames(dataset02)[1] <- "Subject" # first column name gets messed up for some reason
dataset03 <- left_join(dataset02,dataset01[,c(1,4:29)],by="Subject") # merge datasets

library(ggplot2)
```

## Outline

- Background

- Data Management

- Exploratory Data Analysis

- Main Aims and Preliminary Models

- Future Work

## Background

- With an aging population, car companies must keep the interests of the elderly in mind.

- One particular issue they face is simply getting in and out of the vehicle.

## Background

- Dr. Galecki provided two specific aims for us to answer, but I also wanted to focus my analysis towards results we can implement.

# Data Management

## Data Management

1. Discarded shoe size variable completely from the first dataset.

## Data Management

2. Took the **peak performance** of all repeated measurements

    - Knee Extension Strength
    - Hip Abduction Strength
    - Hip Abduction Rate of Torque
    - Grip Strength
    - Timed Up and Go
    - One Leg Balancing

## Data Management

3. Kept track of number of missing measurements for each subject

    - Not worried about One Leg Balancing Measurements

## Data Management

4. Created categorical One Leg Balancing variable

    - Sum of peak performance in each leg
    - Poor balance if less than 15
    
## Data Management

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
hist(dataset03$OLB_R + dataset03$OLB_L, main = "Distribution of Sum Times", xlab = "Sum of Max Times from Each Leg", ylab = "Frequency", cex.lab=1.5, cex.axis=1.5, cex.main=2, cex.sub=1.5)
```

## Data Management

5. Analyzed correlation structures in each of the following categories to find a single representative variable

    - Anthropometric
    - Strength
    - Mobility

## Data Management

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(corrplot)
body.m <- cor(dataset03[,c(3,5:18)])
corrplot(body.m, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

# Exploratory Data Analysis

## EDA

```{r echo=FALSE, message=FALSE, warning=FALSE}
facet_labels <- c(
  '0' = "Parking Lot",
  '1' = "Open"
)
plot01 <- ggplot(data = dataset03,
                 aes(y = ingress, x = vehicle, group = vehicle)) +
  geom_boxplot() +
  facet_grid(.~open, labeller = as_labeller(facet_labels)) +
  ggtitle("Ingress Time by each Vehicle/Door Position") +
  xlab("Vehicle") +
  ylab("Ingress Time (seconds)") +
  theme_bw() +
  theme(text = element_text(size = 28),
        plot.title = element_text(hjust = 0.5))
plot01
```

## EDA

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
plot04 <- ggplot(data = dataset03,
                 aes(x = vehicle_door, y = egress, group = Subject, color = Group)) +
  geom_line() +
  ggtitle("Egress Time by Subject") +
  xlab("Vehicle-Door Combination") +
  ylab("Egress Time (seconds)") +
  theme_bw() +
  theme(text = element_text(size = 28),
        plot.title = element_text(hjust = 0.5))
plot04
```

## EDA Summary

- Random effects appear to be appropriate for both vehicles and subjects

- What makes Vehicle #2 and Vehicle #7 different?

# Main Aims and Preliminary Models

## Aim #1

### Does egress (ingress) strategy affect driver egress (ingress) time in the elderly?

### Do mobility impairments affect driver egress (ingress) time in the elderly?

## Aim #2

### Identify the human capacities that best explain the variation in egress (ingress) time between drivers.

## Preliminary Models

- Linear model adjusting for age, gender, vehicle, door position, weight, hip strength, poor balance, group, and strategy

    - ingress strategy nearly significant (p=0.06)
    - egress strategy nearly significant (p=0.06)

## Preliminary Models

- Linear mixed model with random intercepts for vehicle and subject

    - ingress strategy highly significant (p<0.01)
    - egress strategy not significant (p=0.18)
    
## Preliminary Models

- Linear mixed model with random intercept for vehicle and random slope for door position by vehicle

    - Resulted in a singular fit
    - Too few random effect levels (7)
    - Indicates overfitting the model

# Future Work

## Future Work

- Identify the human capacities that best explain variation in ingress/egress times (Aim #2)

## Future Work

- Identify which dimensions (if any) have a significant effect on ingress/egress times

    - An explanation to why drivers did better with certain vehicles
    - Give information to engineers so they can design "elderly-friendly" vehicles

# Questions???